---
title: "tidy-model"
author: "Daniel Teoh"
date: "2025-02-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library('tidyverse');library('modeldata')
```

## Introduction

This R markdown domcument demostrates my abilites to use models for data analysis using a data set collected on crickets.

```{r load data, include = FALSE}
data(crickets, package = "modeldata")
names(crickets)
```

## What is that sound?

In this report, we examine what factors predict a cricket's chirp rate

```{r summary, echo = FALSE}
summary(crickets)
ggplot(crickets , aes(x = rate)) + 
  geom_histogram(bins = 8) +
  ggtitle("Distribution of the chirp rate of crickets") +
  xlab('Chirp rate (per min.)')
```

## Tempreature affects chrip rate
```{r temp, echo= FALSE}
ggplot(crickets, aes(x = temp, y = rate)) +
  geom_point() +
  geom_smooth(method = 'lm') +
  ggtitle("Plot of temperature and chirp rate") +
  ylab('Chirp rate (per min.)') +
  xlab('Temperature (Celsius)')

temp_lm <- lm(rate ~ temp, crickets)

summary.lm(temp_lm)
```
When looking at  the scatter plot of the temperature and chirping correlation
test we can see that the average chirp rate per minute increases at a rate of 
4.2 per minute as the temperature increases 1 degree.