---
title: "tidy-model-diamonds"
author: "Daniel Teoh"
date: "2025-02-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library('tidyverse') ; library('ggplot2')

```

##Introdution 

This R Markdown document demonstrates my ability to build linear models for data analysis using the diamonds dataset from the ggplot2 package. The goal is to predict a diamond's price based on its characteristics.

```{r load data, include = FALSE}
data(diamonds, package = "ggplot2")
names(diamonds)
```

##Distribution of Diamond Prices

In this section we examine the distribution of diamond prices.

```{r summary, echo = FALSE}
summary(diamonds$price)
ggplot(diamonds, aes(x = price)) + 
  geom_histogram(bins = 50, fill = "steelblue", color = "black") +
  ggtitle("Distribution of Diamond Prices") +
  xlab("Price (USD)") +
  ylab("Count")

```

##Carat Weight and Price Relationship

```{r weight , echo = FALSE}
ggplot(diamonds, aes(x = carat, y = price)) +
  geom_point(alpha = 0.2) +
  geom_smooth(method = "lm", color = "red") +
  ggtitle("Carat Weight vs. Price") +
  xlab("Carat") +
  ylab("Price (USD)")

carat_lm <- lm(price ~ carat, data = diamonds)
summary.lm(carat_lm)

```

When looking at the scatter plot of carat weight and price, we can see that the price increases as carat weight increases. The fitted linear model suggests a strong positive correlation between carat and price.

##Multi-Variable Linear Model for Price Prediction
```{r variable, echo = FALSE}
ggplot(diamonds, aes(x = carat, y = price)) +
  geom_point(alpha = 0.2) +
  geom_smooth(method = "lm", color = "red") +
  ggtitle("Carat Weight vs. Price") +
  xlab("Carat") +
  ylab("Price (USD)")

multi_lm <- lm(price ~ carat + cut + color + clarity, data = diamonds)
summary.lm(multi_lm)
```

##Interaction
```{r variable histogram echo = FALSE}
ggplot(diamonds, aes(x = cut, y = price, fill = cut)) +
  geom_boxplot() +
  ggtitle("Effect of Diamond Cut on Price") +
  xlab("Cut") +
  ylab("Price (USD)")
```