---
title: "tidy-model-diamonds"
author: "Daniel Teoh"
date: "2025-02-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library('tidyverse') ; library('ggplot2')

```

##Introdution 

This R Markdown document demonstrates my ability to build linear models for data analysis using the diamonds dataset from the ggplot2 package. The goal is to predict a diamond's price based on its characteristics.

```{r load data, include = FALSE}
data(diamonds, package = "ggplot2")
names(diamonds)
```

##Distribution of Diamond Prices

In this section we examine the distribution of diamond prices.

```{r summary, echo = FALSE}
summary(diamonds$price)
ggplot(diamonds, aes(x = price)) + 
  geom_histogram(bins = 50, fill = "steelblue", color = "black") +
  ggtitle("Distribution of Diamond Prices") +
  xlab("Price (USD)") +
  ylab("Count")

```

##Carat Weight and Price Relationship

```{r weight , echo = FALSE}
ggplot(diamonds, aes(x = carat, y = price)) +
  geom_point(alpha = 0.2) +
  geom_smooth(method = "lm", color = "red") +
  ggtitle("Carat Weight vs. Price") +
  xlab("Carat") +
  ylab("Price (USD)")

carat_lm <- lm(price ~ carat, data = diamonds)
summary.lm(carat_lm)

```

When looking at the scatter plot of carat weight and price, we can see that the price increases as carat weight increases. The fitted linear model suggests a strong positive correlation between carat and price.

##Multi-Variable Linear Model for Price Prediction
```{r variable, echo = FALSE}
ggplot(diamonds, aes(x = carat, y = price)) +
  geom_point(alpha = 0.2) +
  geom_smooth(method = "lm", color = "red") +
  ggtitle("Carat Weight vs. Price") +
  xlab("Carat") +
  ylab("Price (USD)")

multi_lm <- lm(price ~ carat + cut + color + clarity, data = diamonds)
summary.lm(multi_lm)
```
The scatter plot shows a strong positive correlation between carat weight and price, with larger diamonds generally costing more. The red regression line confirms this trend, though the spread suggests other factors also influence price. To account for this, we fit a multi-variable model including cut, color, and clarity, which helps refine price predictions by incorporating these quality attributes. 

##Interaction
```{r variable histogram, echo = FALSE}
ggplot(diamonds, aes(x = cut, y = price, fill = cut)) +
  geom_boxplot() +
  ggtitle("Effect of Diamond Cut on Price") +
  xlab("Cut") +
  ylab("Price (USD)")
```
```{r interation, echo = FALSE}
ggplot(diamonds, aes(x = carat, y = price, color = cut)) + 
  geom_point(alpha = 0.2) +
  geom_smooth(method = "lm") +
  ggtitle("Carat Weight vs. Price Across Different Cuts") +
  xlab("Carat") +
  ylab("Price (USD)")

interaction_lm <- lm(price ~ carat * cut, data = diamonds)
summary.lm(interaction_lm)

anova(multi_lm, interaction_lm)

```